Predicting Infection Outbreak Dynamics in Toronto Healthcare Institutions: A Bayesian Approach

### 2024-12-04 paper/paper.qmd
Update paper.qmd
included appendix

>>>>

# Appendix

# Appendix A: Idealized Survey: Methodology

A trend noticed in our modelling results was that LTCHs report the highest number of outbreaks in comparison to other healthcare settings. It is essential to understand the reasoning behind this in order to better inform policy reform and resource reallocation to curb this issue. Using an idealized survey methodology, this survey aims to identify potential issues in the availability of care, LTCH organizational structure, and other demographic or environmental factors that may influence their pattern of increased outbreaks. Further, it would be ideal to compare the responses of LTCH workers and hospital workers, which reported the least amount of outbreaks overall, to note any discrepancies in operation that may be contributing to the increased outbreaks in LTCH.

Sampling Approach

Population: Our population of interest are LTCH and hospitals that report infectious outbreaks. Specifically, we are aiming to gain information about the nature of work settings in these healthcare institutions. For this, it is best to contact hospital and LTCH workers, administrators and nursing staff to understand the quality of care provided to the patients as well as the quality of infection control and disease prevention employed at these institutions. A comparison of the nature of infection control practices conducted at hospitals and LTCH will also provide key insights into the issue. 
Frame: The frame includes workers of LTCH and hospitals in Toronto who have worked in this setting for the past year. This frame can be obtained from hospital and LTCH people directories, by directly contacting these healthcare institutions, and involving Toronto Public Health in the data collection process. 
Sample: The sample for this survey would be a subset of all hospital and LTCH workers. The sample size must be sufficient enough to make the results of the survey viable for further statistical analysis. We also have to ensure that all aspects of the outbreaks are covered by this sampling frame, which can be done by including specific questions about the nature of different outbreaks as influenced by the time of year, causative agents, and type of care received. 
This survey follows a stratified random sampling approach to ensure that all types of healthcare settings and causative agents are captured in the data collected. Further, we also want to ensure that a vast array of participants within the healthcare institution care network are represented - such as nurses, administrators, on-call and permanent doctors, critical care unit staff, specific staff for the three types of hospital settings (acute, chronic and psychiatric), etc. This would include a wide array of healthcare workers and allow diversification of opinions, allowing for a holistic understanding of outbreak prevention and management practices in these institutions. 
Recruitment Strategy

Given that we wish to collect data on specific healthcare institutions instead of individual caregivers, it would be apt to target healthcare institutions to be able to reach all their staff. This can be done by contacting the list of institutions listed in our original Open Data Toronto dataset, contacting other listed hospitals and LTCH as listed by Toronto Public Health, and also contacting any community-based facilities that may operate in the city. These institutions can be contacted by email, telephone, etc. Any individual healthcare workers who aren’t actively working in a healthcare setting can also be contacted by reviewing the healthcare worker directory by Toronto Public Health and being contacted via email or phone, or by running ads in Toronto (in subway stations, on the radio, on social media, at universities) requesting healthcare workers to fill out our survey. This ensures that we not only capture active healthcare staff but also staff that may be on leave, may have worked in healthcare settings in the past, or may have switched careers or healthcare institutions.

Non-Response Handling

If healthcare institutions do not respond, we can follow up with them via alternative modes of communication such as emailing other departments at the institution, follow-up phone calls, contacting them via social media, etc. Individual healthcare workers can also be provided additional incentives such as giveaways or a remuneration for completing the survey. Workers will also be ensured that all their data will be anonymous to promote participation. Further, healthcare institutions will be briefed upon the reason for this study, the potential for resource allocation to alleviate the stress on LTCH and to better support them, and to curb the level of infection outbreaks in Toronto. These reasons will promote healthcare institutions to participate in the study - for reduced outbreak levels and better funding opportunities.

Strengths & Limitations

This survey aims to gain specific insight regarding the trend of high infection outbreaks in LTCH as compared to other healthcare institutions, specifically hospitals, that have reported a significantly smaller number of outbreaks in comparison to LTCH. This survey thus targets healthcare workers, caregivers, staff and administrators to understand the infection prevention and control practices at these institutions as well as the quality of care given. 
This survey is a robust tool for understanding the granular, facility-level difficulties an institution might undergo in handling outbreaks. By collecting institution-by-institution data, we can ensure that resource reallocation, budget reforms, and overall policy reforms can be made at the district-level, and not just the city-wide level to meet the individual needs of institutions or a group of institutions. A comparison between LTCH and hospital care practices also provides us with specific insight into how to better handle certain prevention procedures. Further, information about specific institutions can allow us to identify potential heatmaps in the city where greater outbreaks occur, and understand any underlying socio-economic or environmental factors contributing to this. For example, LTCH may quote lower staff-to-resident ratios, older building infrastructure or shared dining/recreational areas as reasons for higher outbreaks. Further, LTCHs tend to have older residents that may have comorbid conditions and increased health vulnerabilities due to a weaker immune system.

However, this survey has potential limitations as well. As we do not offer any major incentives for filling the survey, we may suffer from a lower participation rate. Many healthcare institutions that do not have high outbreak levels will be unmotivated to participate in this study, specifically if they fear that resource allocation by Toronto Public Health may be an option. This survey relies on self-reports by the workers regarding their institution’s best practices, which can face issues with compliance and high reporting inaccuracy. Further, if healthcare workers are unhappy at their workplace, they are more likely to bias the results. To tackle this issue, the survey includes a few questions to gauge healthcare worker stress levels and satisfaction levels. Lastly, these results are localised for Toronto LTCH and hospitals and may not be generalizable outside the city. 
By conducting this survey, we aim to understand the reasoning behind higher outbreaks in LTCH as compared to hospitals in Toronto. This informs healthcare institutions, policymakers and healthcare workers of the best practices regarding outbreak prevention and furthers the dialogue for patient safety, alleviating healthcare worker stress, tackling staffing issues, infrastructure developments, and better resource allocation. 
Idealized Survey

A sample survey was created using Google Forms, and can be found here: https://forms.gle/aQzZHHpW8sjM9Vog8

All survey questions can be found below as well.

Toronto Infection Outbreak Tracking Survey

Introduction

Hello! Thank you so much for participating in this study. We aim to understand the nature of infection outbreak prevention and control strategies employed in Toronto’s various healthcare institutions such as Long-term Care Homes (LTCH) and hospitals. Rest assured that all responses to this survey will be anonymous and confidential. This survey should take approximately 10-15 minutes to complete.

If you have any questions about the survey or the research study, wish to learn more about it, or wish to withdraw from the study, please contact the Principal Investigator:

Name: Divya Gupta
Email: div.gupta@utoronto.ca

Thank you!

Section 1: Demographic Information

What is your age?
18-20
20-30
30-40
40-60
60+
What is your gender?
Female
Male
Non-binary
Other:
What is your role in the healthcare facility?
Nurse
Doctor
Administrator
Caregiver
Cleaning Staff
Other:
What healthcare facility type do you work for?
LTCH
Hospital - Acute Care
Hospital - Chronic Care
Hospital - Psychiatric Care
Other:
What neighborhood of Toronto do you work in?
How many years have you worked in a healthcare setting?  
 less than a year
1-3 years
3-5 years
6-10 years
10+ years
How many years of experience do you have with infection outbreak control and prevention practices in healthcare settings?
less than a year
1-3 years
3-5 years
6-10 years
10+ years
 Have you received formal training on infection outbreak prevention and control?
 Yes, this past year
Yes, 1-2 years ago
Yes, more than 2 years ago
Yes, when I was onboarded
No
Do you hold any professional certification related to infection control or healthcare management?
Yes
No
How often have you been directly involved in handling outbreaks at your facility in the past 12 months?
never
once or twice
3-5 times
multiple times

Section 2: Facility Practices
How frequently are common areas cleaned and disinfected?

Multiple times daily
Daily
Weekly
Less often
Are there isolation protocols for residents exhibiting symptoms of illness?
Yes
No

What infection control measures are currently implemented? (select all that apply)
Use of personal protective equipment (PPE)
Hand hygiene stations in key areas
Regular testing of residents and staff
Temperature checks at facility entrances

How often are HVAC (Heating, Ventilation, and Air Conditioning) systems inspected?
Monthly
Every 6 months
Annually
Never
Not sure

How often are infection prevention training sessions conducted for staff?
Monthly
Quarterly
Annually
Never

Are there dedicated staff for cleaning and sanitation?
Yes
No
Section 3: Staffing
What is the average staff-to-resident ratio during different shifts?

Morning shift: ___
Afternoon shift: ___
Night shift: ___
Have staffing shortages occurred in the past year?

Yes, frequently
Yes, occasionally
No
How often do staff rotate between different roles (e.g., caregiving, cleaning)?

Frequently
Occasionally
Rarely
Never
Are staff provided with paid sick leave?

Yes
No
Do you feel adequately supported in managing outbreak-related challenges?
Yes
No
N/A
Section 4: Resident Health and Care
Do residents share rooms?
Yes, all
Yes, some
No, all residents have private rooms

What percentage of residents are immunized annually against:

Influenza: ___ %
COVID-19: ___ %
Norovirus: ___ %
Are new residents screened for infectious diseases upon admission?

Yes
No
How often are residents tested for respiratory and gastrointestinal infections during outbreak seasons?

Weekly
Bi-weekly
Monthly
Only during visible symptoms

How often are residents tested for COVID-19?
Weekly
Bi-weekly
Monthly
Only during visible symptoms

Do residents regularly participate in group activities?

Yes, frequently
Yes, occasionally
No
Are residents provided with education about hygiene practices (e.g., handwashing)?

Yes
No
Are dining and recreation areas shared among residents?
Yes
No
Does the facility have outdoor areas where residents can spend time?

Yes
No
Are family members of residents informed promptly during outbreaks?
Yes
No
Section 5: Emergency Preparedness
Does the facility have a written outbreak management plan?

Yes
No
How frequently are outbreak response drills conducted?

Monthly
Quarterly
Annually
Never

Do you find that the outbreak response drills are useful for training?
Yes, very 
Yes, somewhat
Neutral
No, never attended
No, not useful
Other:

What is the time frame for notifying public health authorities about a suspected outbreak?

Within 24 hours
Within 48 hours
Within 3 business days
Longer

Is additional staffing or support brought in during outbreaks?
Yes
No

What challenges does your facility face in implementing outbreak prevention and control measures?

Are there any specific recommendations you would like to make for improving infection control in long-term care homes?
Final Section
Thank you so much for participating in this survey! Feel free to contact us by emailing div.gupta@utoronto.ca in case of any questions.
